---
title: "MinneMUDAC Data Analytics Competition Visualizations"
author: "Bjorn Melin"
date: "11/7/2019"
output: html_document
---

## Questions To Answer With Data:  
1.  What trends do you see in U.S. soybean exports?
2.  What seasonality do you see in the data?
3.  Do you see any relationships between commodity exports (substitutable products: corn/soybeans)?
4.  Are we trade neutral for agricultural products in a given month/year (consider columns past BE)?
5.  Does competitor exports of soybeans (column AK) impact US exports of soybeans?

\  
\  
\  
\   


## Load In Tidy Datasets W/ Only Highly Correlated Features To Soybeans:

```{r}
# Load the dataset
us_competitor_top_corr <- read.csv("us_competitor_top_corr.csv", stringsAsFactors = TRUE)
us_suppliers_top_corr <- read.csv("us_suppliers_top_corr.csv", stringsAsFactors = TRUE)
us_markets_top_corr <- read.csv("us_markets_top_corr.csv", stringsAsFactors = TRUE)

subset(ChickWeight, Diet==4 && Time == 21)

us_competitor_top_corr
us_suppliers_top_corr
us_markets_top_corr
```  


## Modify Datasets To Only Include Data Since 2005:  

```{r}
us_competitor_subset <- subset(us_competitor_top_corr, Year>=2010) %>% 
  unite("Date", Month:Year, sep="/", remove=TRUE)
us_suppliers_subset <- subset(us_suppliers_top_corr, Year>=2010) %>% 
  unite("Date", Month:Year, sep="/", remove=TRUE)
us_markets_subset <- subset(us_markets_top_corr, Year>=2010) %>% 
  unite("Date", Month:Year, sep="/", remove=TRUE)

us_competitor_subset
us_suppliers_subset
us_markets_subset


write.csv(us_competitor_subset, "us_competitor_subset.csv", row.names = FALSE)
write.csv(us_suppliers_subset, "us_suppliers_subset.csv", row.names = FALSE)
write.csv(us_markets_subset, "us_markets_subset.csv", row.names = FALSE)
```




## Trends Seen In US Soybean Exports Since 2005:

```{r}
library(tidyverse)
library(ggplot2)

plot(us_markets_subset$Soybeans, type="l", xlab="Date", ylab="Soybean Exchange Rate",
     main="US Market Soybean Exchange Rates (2010-2019)")
```


## Trends Seen In US Soybean Exports Since 1970:


```{r}
library(tidyverse)
library(ggplot2)

us_markets_date <- us_markets_top_corr %>% 
  unite("Date", Month:Year, sep="/", remove=TRUE)

plot(us_markets_date$Soybeans, type="l", xlab="Date", ylab="Soybean Exchange Rate",
     main="US Market Soybean Exchange Rates (1970-2019)")
```



## Seasonality In US Markets Data:


## Seasonality In US Competitor Data:


## Seasonality In US Markets Data:



