---
title: "MinneMUDAC Data Analytics Competition"
author: "Bjorn Melin"
date: "10/24/19"
output:
  html_document:
    df_print: paged
---

## Load All Libraries Used:  

```{r, message=FALSE, warning=FALSE}
# Load the libraries we will need
library(Cubist)
library(psych)
library(tidyverse)
library(ggplot2)
library(directlabels)
```


## Load In Tidy Datasets For US Commodities:

```{r}
# Load the dataset
us_markets <- read.csv("US_Markets.csv", stringsAsFactors = TRUE)
us_competitor <- read.csv("competitor_Markets.csv", stringsAsFactors = TRUE)
us_suppliers <- read.csv("US_Suppliers.csv", stringsAsFactors = TRUE)
# Rename a the Soybeans feature in us_suppliers
colnames(us_suppliers)[colnames(us_suppliers)=="Oils..ex..Soybean." ] <- "Soybeans"

us_markets %>% head(4)
us_suppliers %>% head(4)
us_competitor %>% head(4)
```  



## Run `Cubist` Algorithm To Find Features With High Correlation To Soybeans For US Markets:

```{r, message=FALSE, warning=FALSE}
# Tidy the us_markets dataset by removing variables which will not be of interest
us_markets_tidy <- us_markets[-c(1:4)]
#us_markets_tidy
#str(us_markets_tidy)

#pairs.panels(us_markets_tidy[1:25])
#cor.plot(us_markets_tidy, MAR = TRUE, symmetric = TRUE)

# Create the training and 
sample_exchange_markets_train <- us_markets_tidy[1:600, ]

# Run Cubist Algorithm On Training and Testing Datasets to predict Soybean
exchange_model_markets <- cubist(sample_exchange_markets_train[-5], sample_exchange_markets_train$Soybeans)
#exchange_model_markets
summary(exchange_model_markets)
```




## Run `Cubist` Algorithm To Find Features With High Correlation To Soybeans For US Suppliers:

```{r}
# Tidy the us_suppliers dataset by removing variables which will not be of interest
us_suppliers_tidy <- us_suppliers[-c(1:3)]
#(us_suppliers_tidy)

# Create the training and 
sample_exchange_suppliers_train <- us_suppliers_tidy[1:600, ]

# Run Cubist Algorithm On Training and Testing Datasets to predict Soybean
exchange_model_suppliers <- cubist(sample_exchange_suppliers_train[-7], sample_exchange_suppliers_train$Soybeans)
#exchange_model_suppliers
summary(exchange_model_suppliers)
```






## Run `Cubist` Algorithm To Find Features With High Correlation To Soybeans For US Competitors:

```{r}
# Tidy the us_competitor dataset by removing variables which will not be of interest
us_competitor_tidy <- us_competitor[-c(1:4)]
#us_competitor_tidy

# Create the training and 
sample_exchange_comp_train <- us_competitor_tidy[1:600, ]

# Run Cubist Algorithm On Training and Testing Datasets to predict Soybean
exchange_model_comp <- cubist(sample_exchange_comp_train[-5], sample_exchange_comp_train$Soybeans)
#exchange_model_comp
summary(exchange_model_comp)
```



## Remove All Vars Below 90% Correlated To Soybeans From US Markets DF:

```{r}
# Create a new tidied data frame only containing 
# variables which are above 90% correlated with Soybeans
us_markets_top_corr <- us_markets[c("Month", "Year", "U.S..merchandise.exports.to.the.World",   
                                    "U.S..agricultural.exports.to.the.World", "Soybeans", "Hides...Skins", 
                                    "Wheat", "Corn", "Bulk", "Cotton", "Wine...Beer")]
#us_markets_top_corr
write.csv(us_markets_top_corr, "us_markets_top_corr.csv", row.names = FALSE)
```


## Remove All Vars Below 90% Correlated To Soybeans From Suppliers DF:

```{r}
# Create a new tidied data frame only containing 
# variables which are above 90% correlated with Soybeans
us_suppliers_top_corr <- us_suppliers[c("Month", "Year", "U.S..agricultural.imports",   
                                     "Soybeans", "Intermediates", "Fish.Products", 
                                     "Cocoa.Paste....Butter", "Coffee..Unroasted",
                                     "Chocolate...Cocoa.Products")]
#us_suppliers_top_corr
write.csv(us_suppliers_top_corr, "us_suppliers_top_corr.csv", row.names = FALSE)
```



## Remove All Vars Below 90% Correlated To Soybeans From US competitors DF:

```{r}
# Create a new tidied data frame only containing 
# variables which are above 90% correlated with Soybeans
us_competitor_top_corr <- us_competitor[c("Month", "Year", "Competitor.merchandise.exports.to.the.World" ,   
                                    "Competitor.agricultural.exports.to.the.World", "Soybeans", 
                                    "Tobacco", "Corn", "Soymeal", "Bulk", "Rice", "Tree.Nuts",
                                    "Dairy.Products", "Fresh...Processed.Fruit",
                                    "Fresh...Processed.Vegetables", "Beef...Beef.Products",
                                    "Wine...Beer")]
#us_competitor_top_corr
write.csv(us_competitor_top_corr, "us_competitor_top_corr.csv", row.names = FALSE)
```



```{r}
# x is a matrix containing the data
# method : correlation method. "pearson"" or "spearman"" is supported
# removeTriangle : remove upper or lower triangle
# results :  if "html" or "latex"
  # the results will be displayed in html or latex format
corstars <-function(x, method=c("pearson", "spearman"), removeTriangle=c("upper", "lower"),
                     result=c("none", "html", "latex")){
    #Compute correlation matrix
    require(Hmisc)
    x <- as.matrix(x)
    correlation_matrix<-rcorr(x, type=method[1])
    R <- correlation_matrix$r # Matrix of correlation coeficients
    p <- correlation_matrix$P # Matrix of p-value 
    
    ## Define notions for significance levels; spacing is important.
    mystars <- ifelse(p < .0001, "****", ifelse(p < .001, "*** ", ifelse(p < .01, "**  ", ifelse(p < .05, "*   ", "    "))))
    
    ## trunctuate the correlation matrix to two decimal
    R <- format(round(cbind(rep(-1.11, ncol(x)), R), 2))[,-1]
    
    ## build a new matrix that includes the correlations with their apropriate stars
    Rnew <- matrix(paste(R, mystars, sep=""), ncol=ncol(x))
    diag(Rnew) <- paste(diag(R), " ", sep="")
    rownames(Rnew) <- colnames(x)
    colnames(Rnew) <- paste(colnames(x), "", sep="")
    
    ## remove upper triangle of correlation matrix
    if(removeTriangle[1]=="upper"){
      Rnew <- as.matrix(Rnew)
      Rnew[upper.tri(Rnew, diag = TRUE)] <- ""
      Rnew <- as.data.frame(Rnew)
    }
    
    ## remove lower triangle of correlation matrix
    else if(removeTriangle[1]=="lower"){
      Rnew <- as.matrix(Rnew)
      Rnew[lower.tri(Rnew, diag = TRUE)] <- ""
      Rnew <- as.data.frame(Rnew)
    }
    
    ## remove last column and return the correlation matrix
    Rnew <- cbind(Rnew[1:length(Rnew)-1])
    if (result[1]=="none") return(Rnew)
    else{
      if(result[1]=="html") print(xtable(Rnew), type="html")
      else print(xtable(Rnew), type="latex") 
    }
} 
```


```{r}
us_markets_tidy_cor <- round(cor(us_markets_tidy),2)
# us_markets_tidy_cor

cor_markets_plot <- corstars(us_markets_tidy_cor)
cor_markets_plot
```




# Creating Data Visualizations From Tidy Data:  

## Modify Datasets To Only Include Data Since 2010:  

```{r}
us_competitor_subset <- subset(us_competitor_top_corr, Year>=2010) %>% 
  unite("Date", Month:Year, sep="/", remove=TRUE)
us_suppliers_subset <- subset(us_suppliers_top_corr, Year>=2010) %>% 
  unite("Date", Month:Year, sep="/", remove=TRUE)
us_markets_subset <- subset(us_markets_top_corr, Year>=2010) %>% 
  unite("Date", Month:Year, sep="/", remove=TRUE)

#us_competitor_subset
#us_suppliers_subset
#us_markets_subset


write.csv(us_competitor_subset, "us_competitor_subset.csv", row.names = FALSE)
write.csv(us_suppliers_subset, "us_suppliers_subset.csv", row.names = FALSE)
write.csv(us_markets_subset, "us_markets_subset.csv", row.names = FALSE)
```




## Trends Seen In US Soybean Exports Since 2010:

```{r}
plot(us_markets_subset$Soybeans, type="l", xlab="Months Since Jan 2010", ylab="Soybean Exchange Rate",
     main="US Market Soybean Exchange Rates (2010-2019)")

```


## Trends Seen In US Soybean Exports Since 1970:


```{r}
us_markets_date <- us_markets_top_corr %>% 
  unite("Date", Month:Year, sep="/", remove=TRUE)

plot(us_markets_date$Soybeans, type="l", xlab="Months Since Jan 1970", ylab="Soybean Exchange Rate",
     main="US Market Soybean Exchange Rates (1970-2019)")

#ggplot(data=us_markets_top_corr, aes(x=Year, y=Soybeans, group=Month)) +
#  geom_line(color=us_markets_top_corr$Month) 
```


## Competitor Exports Of Soybeans vs. US Exports Of Soybeans Since 1970:  


```{r}
us_vs_competitor_temp <- cbind(us_markets_top_corr, us_competitor_top_corr$Soybeans)

# Rename a the Soybeans feature in us_vs_competitor_temp
colnames(us_vs_competitor_temp)[colnames(us_vs_competitor_temp)=="Soybeans" ] <- "US_Soybeans"
# Rename a the Soybeans feature in us_vs_competitor_temp
colnames(us_vs_competitor_temp)[colnames(us_vs_competitor_temp)=="us_competitor_top_corr$Soybeans" ] <- "Competitor_Soybeans"

us_vs_competitor <- us_vs_competitor_temp[c("Month", "Year", "US_Soybeans", "Competitor_Soybeans")]

#us_vs_competitor

us_vs_competitor_dif <- cbind(us_vs_competitor, (us_vs_competitor$US_Soybeans)- (us_vs_competitor$Competitor_Soybeans))

plot(us_vs_competitor_dif$`(us_vs_competitor$US_Soybeans) - (us_vs_competitor$Competitor_Soybeans)`, type="l", )
```


## Competitor Exports Of Soybeans vs. US Exports Of Soybeans Since 2010:  


```{r}
us_vs_competitor_temp_subset <- cbind(us_markets_top_corr, us_competitor_top_corr$Soybeans) %>% 
  subset(Year >= 2010)
#us_vs_competitor_temp_subset

# Rename a the Soybeans feature in us_vs_competitor_temp
colnames(us_vs_competitor_temp_subset)[colnames(us_vs_competitor_temp_subset)=="Soybeans" ] <- "US_Soybeans"
# Rename a the Soybeans feature in us_vs_competitor_temp
colnames(us_vs_competitor_temp_subset)[colnames(us_vs_competitor_temp_subset)=="us_competitor_top_corr$Soybeans" ] <- "Competitor_Soybeans"

us_vs_competitor_subset <- us_vs_competitor_temp_subset[c("Month", "Year", "US_Soybeans", "Competitor_Soybeans")]

#us_vs_competitor_subset

us_vs_competitor_subset_dif <- cbind(us_vs_competitor_subset, (us_vs_competitor_subset$US_Soybeans)- (us_vs_competitor_subset$Competitor_Soybeans))

plot(us_vs_competitor_subset_dif$`(us_vs_competitor_subset$US_Soybeans) - (us_vs_competitor_subset$Competitor_Soybeans)`, type="l", )
```


## Seasonality In US Markets Data:

```{r}
us_markets_subset_seasons <- subset(us_markets_top_corr, Year>=2010) %>% 
    select(1,2,5)
us_markets_subset_seasons$Month = as.numeric(as.character(us_markets_subset_seasons$Month))
us_markets_subset_seasons$Year = as.numeric(as.character(us_markets_subset_seasons$Year))

us_markets_subset_seasons <- us_markets_subset_seasons %>% 
    mutate(
    Season = case_when(
      Month %in% 9:11 ~ "Fall",
      Month %in%  c(12,1:2)  ~ "Winter",
      Month %in%  3:5  ~ "Spring",
      TRUE ~ "Summer"))

#us_markets_subset_seasons

write.csv(us_markets_subset_seasons, "us_markets_subset_seasons.csv", row.names = FALSE)


plot <- ggplot(data=us_markets_subset_seasons, aes(x=Month, y=Soybeans, group=Year, colour=Year)) +
  geom_line() +
  xlim(0,13) +
  ylim(90.5,110) +
  labs(xlab="Month", ylab="Soybeans Exchange Rate", 
       title="Monthly Soybean Exhange Rate Trends - US Markets") +
  theme(plot.title = element_text(hjust = 0.5))

print(direct.label(plot,list("last.qp",rot=30)))
```


## Seasonality In US Competitor Data:

```{r}
us_competitor_subset_seasons <- subset(us_competitor_top_corr, Year>=2010) %>% 
  select(1,2,5)
us_competitor_subset_seasons$Month = as.numeric(as.character(us_competitor_subset_seasons$Month))
us_competitor_subset_seasons$Year = as.numeric(as.character(us_competitor_subset_seasons$Year))

us_competitor_subset_seasons <- us_competitor_subset_seasons %>% 
    mutate(
    Season = case_when(
      Month %in% 9:11 ~ "Fall",
      Month %in%  c(12,1:2)  ~ "Winter",
      Month %in%  3:5  ~ "Spring",
      TRUE ~ "Summer"))

#us_competitor_subset_seasons

write.csv(us_competitor_subset_seasons, "us_competitor_subset_seasons.csv", row.names = FALSE)

plot <- ggplot(data=us_competitor_subset_seasons, aes(x=Month, y=Soybeans, group=Year, colour=Year)) +
  geom_line() +
  xlim(0,13) +
  ylim(90.5,130) +
  labs(xlab="Month", ylab="Soybeans Exchange Rate", 
       title="Monthly Soybean Exhange Rate Trends - US Competitors") +
  theme(plot.title = element_text(hjust = 0.5))

print(direct.label(plot,list("last.qp",rot=30)))
```


## Seasonality In US Suppliers Data:

```{r}
us_suppliers_subset_seasons <- subset(us_suppliers_top_corr, Year>=2010) %>% 
  select(1,2,4)
us_suppliers_subset_seasons$Month = as.numeric(as.character(us_suppliers_subset_seasons$Month))
us_suppliers_subset_seasons$Year = as.numeric(as.character(us_suppliers_subset_seasons$Year))

us_suppliers_subset_seasons <- us_suppliers_subset_seasons %>% 
    mutate(
    Season = case_when(
      Month %in% 9:11 ~ "Fall",
      Month %in%  c(12,1:2)  ~ "Winter",
      Month %in%  3:5  ~ "Spring",
      TRUE ~ "Summer"))

#us_suppliers_subset_seasons

write.csv(us_suppliers_subset_seasons, "us_suppliers_subset_seasons.csv", row.names = FALSE)

plot <- ggplot(data=us_suppliers_subset_seasons, aes(x=Month, y=Soybeans, group=Year, colour=Year)) +
  geom_line() +
  xlim(0,13) +
  ylim(90.5,130) +
  labs(xlab="Month", ylab="Soybeans Exchange Rate", 
       title="Monthly Soybean Exhange Rate Trends - US Suppliers") +
  theme(plot.title = element_text(hjust = 0.5))

print(direct.label(plot,list("last.qp",rot=30)))
```




