---
title: "MinneMUDAC Data Analytics Competition"
author: "Bjorn Melin"
date: "10/24/19"
output:
  html_document:
    df_print: paged
---



## Load In Tidy Datasets For US Commodities:

```{r}
# Load the dataset
us_markets <- read.csv("US_Markets.csv", stringsAsFactors = TRUE)
us_suppliers <- read.csv("US_Suppliers.csv", stringsAsFactors = TRUE)
# Rename a the Soybeans feature in us_suppliers
colnames(us_suppliers)[colnames(us_suppliers)=="Oils..ex..Soybean." ] <- "Soybeans"

us_competitor <- read.csv("competitor_Markets.csv", stringsAsFactors = TRUE)
#us_markets
#us_suppliers
#us_competitor
#us_markets
```  



## Run `Cubist` Algorithm To Find Features With High Correlation To Soybeans For US Markets:

```{r}
# Load the libraries we will need
library(Cubist)
library(psych)

# Tidy the us_markets dataset by removing variables which will not be of interest
us_markets_tidy <- us_markets[-c(1:4)]
#us_markets_tidy
str(us_markets_tidy)

#pairs.panels(us_markets_tidy[1:25])
#cor.plot(us_markets_tidy, MAR = TRUE, symmetric = TRUE)

# Create the training and 
sample_exchange_markets_train <- us_markets_tidy[1:600, ]

# Run Cubist Algorithm On Training and Testing Datasets to predict Soybean
exchange_model_markets <- cubist(sample_exchange_markets_train[-5], sample_exchange_markets_train$Soybeans)
exchange_model_markets
summary(exchange_model_markets)
model_pred_markets <- predict(exchange_model_markets, us_markets_tidy)
cor(us_markets_tidy$Soybeans, model_pred_markets)
```




## Run `Cubist` Algorithm To Find Features With High Correlation To Soybeans For US Suppliers:

```{r}
# Tidy the us_suppliers dataset by removing variables which will not be of interest
us_suppliers_tidy <- us_suppliers[-c(1:3)]
#(us_suppliers_tidy)

# Create the training and 
sample_exchange_suppliers_train <- us_suppliers_tidy[1:600, ]

# Run Cubist Algorithm On Training and Testing Datasets to predict Soybean
exchange_model_suppliers <- cubist(sample_exchange_suppliers_train[-7], sample_exchange_suppliers_train$Soybeans)
exchange_model_suppliers
summary(exchange_model_suppliers)
model_pred_suppliers <- predict(exchange_model_suppliers, us_suppliers_tidy)
cor(us_suppliers_tidy$Soybeans, model_pred_suppliers)
```






## Run `Cubist` Algorithm To Find Features With High Correlation To Soybeans For US Competitors:

```{r}
# Tidy the us_competitor dataset by removing variables which will not be of interest
us_competitor_tidy <- us_competitor[-c(1:4)]
us_competitor_tidy

# Create the training and 
sample_exchange_comp_train <- us_competitor_tidy[1:600, ]

# Run Cubist Algorithm On Training and Testing Datasets to predict Soybean
exchange_model_comp <- cubist(sample_exchange_comp_train[-5], sample_exchange_comp_train$Soybeans)
exchange_model_comp
summary(exchange_model_comp)
model_pred_comp <- predict(exchange_model_comp, us_competitor_tidy)
cor(us_competitor_tidy$Soybeans, model_pred_comp)
```



#### Remove All Vars Below 90% Correlated To Soybeans From US Markets DF:

```{r}
# Create a new tidied data frame only containing 
# variables which are above 90% correlated with Soybeans
us_markets_top_corr <- us_markets[c("Month", "Year", "U.S..merchandise.exports.to.the.World",   
                                    "U.S..agricultural.exports.to.the.World", "Soybeans", "Hides...Skins", 
                                    "Wheat", "Corn", "Bulk", "Cotton", "Wine...Beer")]
us_markets_top_corr
```


## Remove All Vars Below 90% Correlated To Soybeans From Suppliers DF:

```{r}
# Create a new tidied data frame only containing 
# variables which are above 90% correlated with Soybeans
us_suppliers_top_corr <- us_suppliers[c("Month", "Year", "U.S..agricultural.imports",   
                                     "Soybeans", "Intermediates", "Fish.Products", 
                                     "Cocoa.Paste....Butter", "Coffee..Unroasted",
                                     "Chocolate...Cocoa.Products")]
us_suppliers_top_corr
```



## Remove All Vars Below 90% Correlated To Soybeans From US competitors DF:

```{r}
# Create a new tidied data frame only containing 
# variables which are above 90% correlated with Soybeans
us_competitor_top_corr <- us_competitor[c("Month", "Year", "Competitor.merchandise.exports.to.the.World" ,   
                                    "Competitor.agricultural.exports.to.the.World", "Soybeans", 
                                    "Tobacco", "Corn", "Soymeal", "Bulk", "Rice", "Tree.Nuts",
                                    "Dairy.Products", "Fresh...Processed.Fruit",
                                    "Fresh...Processed.Vegetables", "Beef...Beef.Products",
                                    "Wine...Beer")]
us_competitor_top_corr
corr <- data.class(us_competitor_top_corr)
corr
```




```{r}
# x is a matrix containing the data
# method : correlation method. "pearson"" or "spearman"" is supported
# removeTriangle : remove upper or lower triangle
# results :  if "html" or "latex"
  # the results will be displayed in html or latex format
corstars <-function(x, method=c("pearson", "spearman"), removeTriangle=c("upper", "lower"),
                     result=c("none", "html", "latex")){
    #Compute correlation matrix
    require(Hmisc)
    x <- as.matrix(x)
    correlation_matrix<-rcorr(x, type=method[1])
    R <- correlation_matrix$r # Matrix of correlation coeficients
    p <- correlation_matrix$P # Matrix of p-value 
    
    ## Define notions for significance levels; spacing is important.
    mystars <- ifelse(p < .0001, "****", ifelse(p < .001, "*** ", ifelse(p < .01, "**  ", ifelse(p < .05, "*   ", "    "))))
    
    ## trunctuate the correlation matrix to two decimal
    R <- format(round(cbind(rep(-1.11, ncol(x)), R), 2))[,-1]
    
    ## build a new matrix that includes the correlations with their apropriate stars
    Rnew <- matrix(paste(R, mystars, sep=""), ncol=ncol(x))
    diag(Rnew) <- paste(diag(R), " ", sep="")
    rownames(Rnew) <- colnames(x)
    colnames(Rnew) <- paste(colnames(x), "", sep="")
    
    ## remove upper triangle of correlation matrix
    if(removeTriangle[1]=="upper"){
      Rnew <- as.matrix(Rnew)
      Rnew[upper.tri(Rnew, diag = TRUE)] <- ""
      Rnew <- as.data.frame(Rnew)
    }
    
    ## remove lower triangle of correlation matrix
    else if(removeTriangle[1]=="lower"){
      Rnew <- as.matrix(Rnew)
      Rnew[lower.tri(Rnew, diag = TRUE)] <- ""
      Rnew <- as.data.frame(Rnew)
    }
    
    ## remove last column and return the correlation matrix
    Rnew <- cbind(Rnew[1:length(Rnew)-1])
    if (result[1]=="none") return(Rnew)
    else{
      if(result[1]=="html") print(xtable(Rnew), type="html")
      else print(xtable(Rnew), type="latex") 
    }
} 
```


```{r}
us_markets_tidy_cor <- round(cor(us_markets_tidy),2)
# us_markets_tidy_cor

cor_markets_plot <- corstars(us_markets_tidy_cor)
cor_markets_plot
```

# Heatmap that displays the correlation amongst the features in us_markets_tidy_cor

```{r}
# heatmap displaying correlation amongst the features of us_markets_tidy_cor

library(reshape2)
mealted_us_markets_tidy_cor <- melt(us_markets_tidy_cor)
head(mealted_us_markets_tidy_cor)
ggheatmap <- ggplot(data = mealted_us_markets_tidy_cor, aes(x=Var1, y=Var2, fill = value)) + geom_tile(color = "white") + scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "US markets correlation") + theme_minimal() + theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1)) + coord_fixed()

## adds the correlation values to the map
ggheatmap + geom_text(aes(Var1, Var2, label = value), color = "black", size = 1.5)
```

# Heatmap that displays the correlation amongst the features in us_competitor_tidy

```{r}
us_competitor_tidy_cor <- round(cor(us_competitor_tidy), 2)

# heatmap displaying correlation amongst the features of us_markets_tidy_cor

library(reshape2)
mealted_us_competitor_tidy_cor <- melt(us_competitor_tidy_cor)
head(mealted_us_competitor_tidy_cor)
ggheatmap_competitor <- ggplot(data = mealted_us_competitor_tidy_cor, aes(x=Var1, y=Var2, fill = value)) + geom_tile(color = "white") + scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "US Competitor correlation") + theme_minimal() + theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1)) + coord_fixed()

## adds the correlation values to the map
ggheatmap_competitor + geom_text(aes(Var1, Var2, label = value), color = "black", size = 1.5)

```

# Heatmap that displays the correlation amongst the features in us_suppliers_tidy

```{r}
us_suppliers_tidy_cor <- round(cor(us_suppliers_tidy), 2)

# heatmap displaying correlation amongst the features of us_markets_tidy_cor

library(reshape2)
mealted_us_suppliers_tidy_cor <- melt(us_suppliers_tidy_cor)
head(mealted_us_suppliers_tidy_cor)
ggheatmap_suppliers <- ggplot(data = mealted_us_suppliers_tidy_cor, aes(x=Var1, y=Var2, fill = value)) + geom_tile(color = "white") + scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "US Competitor correlation") + theme_minimal() + theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1)) + coord_fixed()

## adds the correlation values to the map
ggheatmap_suppliers + geom_text(aes(Var1, Var2, label = value), color = "black", size = 1.5)
```




